Mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$

### 2013-11-19 paper/ms.tex
more editing in the paper

The main purpose of this section is to show how different
observational constraints narrow down the parameters space of allowed
models. Each sub-section presents the effect of adding new pieces of
observational or statistical evidence.

@@

In Figure \ref{fig:halos} we present the results for  the
integrated dark matter halo surface density as a function of the
minimum halo mass $M_{\rm min}$. Each line corresponds to one of the
210 sub-volumes in the Bolshoi simulation. The gray band indicates the
surface density values for LAEs allowed by observations
\citep{Yamada2012}.

This result allows us to understand why only a specific range of
models ${\mathcal M}$ can be expected to be consistent with
observations. From Figure \ref{fig:halos} we can read that models with
a minimum mass $M_{\rm min}>10^{11.5}$\hMsun will always have a
surface number density lower than the observational constrain. This
makes impossible that models in that range can be compatible with
observations, there are simply to few halos compared to observed
LAEs. The opposite is true in models with $M_{\rm
  min}<10^{10.5}$\hMsun that have a surface number density larger
observations. In those cases the occupation fraction $f_{\rm occ}<1.0$
can be tuned in order to lower the halo number density to match
observations.

Figure \ref{fig:halos} also illustrates the impact of cosmic
variance. At fixed mass there is an scatter of $0.3-0.6$ dex in the
number density abundance.  A consequence of this variation is that
models with the same mass range and occupation fraction can have mock
fields with number densities varying up to a factor of $\sim
2-5$. This scatter is comparable to the $0.3$ dex scatter coming from 
observations. In the same Figure \ref{fig:halos} we see that at fixed
number density, around the range allowed by observations, there is an
scatter in the masses of $0.4-0.5$ dex.

This scatter induced by cosmic variance is naturally included in the
mock construction process. This will allows us to have a wider range
of models compatible with observations, as compared with a model that
neglects cosmic variance.

@@

Figure \ref{fig:landscape} presents regions in parameter space $M_{\rm
min}$-$M_{\rm max}$, $M_{\rm min}$-$f_{\rm occ}$ where the KS test yields
values of $P>0.05$ at least for one mock survey, which is the most
conservative assumption for agreement between a model and observations. For those models it
is not possible reject the hypothesis that the simulated and observed
data for the surface number density come from the same parent
distribution. In total, there are between $550$ to $600$ models out of
the original $9000$ models that have at least one mock survey
consistent with observations. By inspection of Figure
\ref{fig:landscape} we see that the halo number abundance is able to
constraint the minimum mass $M_{\rm min}$ to a narrow range, while
$M_{\rm max}$ and $f_{\rm occ}$ remain largely unconstrained.

In Figure \ref{fig:landscape} there are three regions of parameter
space that can be clearly distinguished. The first corresponds to
models where the minimum mass is high $M_{\rm min}>
10^{11.5}\hMsun$. None of these models is compatible with observations
as expected from the arguments presented in the previous
subsection. For these models the number density of LAEs is too low
compared to observations.

The second region corresponds to an intermediate range for the minimum
mass $10^{10.5}\hMsun < M_{\rm min}< 10^{11.5}\hMsun$ where,
regardless of the value of the maximum mass $M_{\rm max}$, it is
possible to tune the occupation fraction $f_{\rm occ}$ to bring some
of the mock surveys into good agreement with observations. In this
region of parameter space one can find two extreme kinds of models.
One kind where the mass interval $\Delta M\equiv \log_{10} M_{\rm max}
- \log_{10}M_{\rm  min}$ is narrow with $\Delta M<1.0$ dex, others
where the mass interval is  extended $\Delta M>1.0$ dex going up to
the maximum halo mass present in the simulation at that redshift, with
$\Delta M = 2.5$ dex in those cases.

@@

Without any additional information our method allows us to infer that
most of the successful models are found in the second and third regions of
parameter space. This result was already expected from halo
abundance calculations shown in Figure \ref{fig:halos} and discussed
in the previous subsection.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.95\linewidth,angle=0]{Fig4_match_P5.pdf}
\end{center} 
\caption{ Number of models with at least $N_{\rm high-P}$ mock surveys
  consistent with the observed surface number density
  distribution in terms of the KS-test values $P>0.05$. Only $\sim
  100$ models have all their mocks consistent with observations. 
  \label{fig:high_success_rate}.}  
\end{figure}

@@

In the previous sub-section we presented a conservative criterion of
agreement by selecting the models that had at least one mock survey
with $P>0.05$ in the KS test. Now we turn to a more strict selection
by requiring all constructed mock surveys to be consistent with
observations.

Figure \ref{fig:high_success_rate} shows the number of models
that have at least $N_{\rm high-P}$ mocks with $P>0.05$. For each
model ${\mathcal M}$ there are 15 different mock surveys.  We discard
the models that have 14 mocks or less consistent with
observations. With this cut we find $\sim 100$ models with all the 15
mock survey realizations with $P>0.05$.  This cut represents a
reduction by a factor of $\sim 6$ with respect to the total number of
models with at least one consistent mock.

Figure \ref{fig:restriction_mock} presents the locii of these models
in the parameter space $M_{\rm min}-M_{\rm max}$ and $M_{\rm
  min}-f_{\rm occ}$. With this constraints the
number of consistent models with  $10.5 < \log_{10}M_{\rm min}< 11.0$ are
greatly reduced. This corresponds to the regions in the parameter
space in Figure \ref{fig:landscape} that already had a low number of
consistent mock surveys. On the other hand, from the right panel in
Figure \ref{fig:restriction_mock} one can see that the reduction of
favored values for the occupation fraction $f_{\rm occ}$ is not strong.

We conclude that conditions on the number density statistics, even if they are
very strict, only put strong constraints on the minimum mass $M_{\rm
  min}$ but not on the other two parameter of the model $M_{\rm max}$
and $f_{\rm occ}$. In the next section we use additional observational
information to constrain these parameters.

@@

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{power_law_correlation_maxden.pdf} 
\hspace{5mm}  
\includegraphics[width=0.46\linewidth,angle=0]{power_law_correlation_meanden.pdf} 
\end{center}
\caption{Values for the free parameters $\theta_{0}$ and $\beta$
in the fitting formula (Eq. \ref{eq:fitting}) for the angular
correlation function. Blue dots correspond to simulations and the
green cross to observations by \citet{Hayashino2004} (left) and
\citet{Ouchi2010} (right). The error bars in the theoretical data correspond
to the quadratic average of the fitting errors for each mock survey
\label{fig:correlation_parameters}}
\end{figure*}

Figure \ref{fig:correlation_parameters} shows the main results in a
$\theta_{0}-\beta$  plane where the average and standard deviation
over the mocks is shown in comparison with the result derived from
observations. In the left panel Figure \ref{fig:correlation_parameters} we see
how the observational ACF measured by \cite{Hayashino2004} 
can be successful in reducing the total number of possible models. The
right panel shows the results for fields in the mock catalog with an
average number density compared against the fit of the data
reported by \cite{Ouchi2010}. In this case we find that there are
large uncertainties coming from the observational data, and only a reduced
fraction of the models can be discarded.

@@

@@

\cite{Hayes2010} showed that the luminosity function for
LAEs at $z=2.2$ is consistent with the escape fraction being constant
for every galaxy regardless of its luminosity. From this results they
derive that almost $90\%$ of the star forming galaxies emit insufficient
Lyman $\alpha$ to be detected, effectively setting the occupation
fraction to be $f_{\rm occ}=0.10$. \cite{Dijkstra2013} use similar
methods to find an average \ly escape fraction in high-z galaxies,
they conclude that there is a range in \ly luminosity where the escape
fraction they derive at $z\sim 3.0$, $f_{\rm esc}=(17\pm 5)\%$, could
be interpreted as an occupation fraction $f_{\rm occ}\sim 0.2$. We
consider a success that our model indeed includes such low occupation
fraction models. The observational evidence from galaxy surveys seem
to disfavor the second family of models with high occupation
fractions. Nevertheless, it remains an open task to explain the narrow
mass range $\Delta M<0.5$ in most of the models.

@@

Observational evidence based on the ACF inferred from photometric
measurements in the Extended Chandra Deep Field South has shown that
the median dark matter masses of h los hosting LAEs is
$\log_{10}M_{\rm  med}=10.9^{+0.5}_{-0.9}$\Msun, with a corresponding
occupation fraction of $1-10\%$  \citep{Gawiser07}.  \cite{Ouchi2010}
presents analysis of LAE observations in the redshift interval
$3.1<z<7.0$ and at $z=3.1$ They quote an average mass for the host
dark matter halos of $M_{h}=2.9^{+24.0}_{-2.9}\times 10^{10}$ \hMsun
with a corresponding duty cycle of $0.008\pm 0.03$.

>>>>

The main purpose of this section is to show how different
observational constraints narrow down the parameters space of allowed
models. Each sub-section presents the effect of adding a new piece of 
observational or statistical evidence.

@@

The right panel in Figure \ref{fig:halos} shows the  integrated dark matter halo surface
density as a function of  minimum halo mass $M_{\rm min}$. Each line
corresponds to one of the 210 sub-volumes in the Bolshoi
simulation. The gray band indicates the surface density values for
LAEs allowed by observations \citep{Yamada2012}. The dashed lines
represent the average values in the fields observed by
\citep{Hayashino2004,Ouchi08}.

This plot allows us to understand why only a specific range of
models ${\mathcal M}$ can be expected to be consistent with
observations. From Figure \ref{fig:halos} we can read that models with
a minimum mass $M_{\rm min}>10^{12}$\hMsun always have a
surface number density lower than the observational constrain, making
them incompatible with observations; there are simply to few halos compared to observed
LAEs. The opposite is true in models with $M_{\rm
  min}<10^{11.0}$\hMsun that have a surface number density larger
observations. In those cases the maximum mass $M_{\rm max}$ and the
occupation fraction $f_{\rm occ}<1.0$  can be tuned in order to lower
the halo number density to match observations.

Figure \ref{fig:halos} also illustrates the impact of cosmic
variance. At fixed minimum mass there is an scatter of $0.3-0.6$ dex
in the number density abundance, which is of the same order of
magnitude as the scatter in the observational data.  As a consequence, the variation in
the number density in mocks for models with the same mass range and occupation
fraction can be by factors of $\sim 2-5$.  This scatter induced by
cosmic variance is naturally included in the mock construction
process. This variance will explain that very different models can be
made compatible with observations.

@@

Figure \ref{fig:landscape} presents regions in parameter space $M_{\rm
min}$-$M_{\rm max}$, $M_{\rm min}$-$f_{\rm occ}$ where the KS test
between the mocks and observations yields values of $P>0.05$ for at
least one mock survey in each model. For those models it
is not possible reject the hypothesis that the simulated and observed
data for the surface number density come from the same parent
distribution.

In total, there are between $550$ to $600$ models out of the original
$9000$ models that have at least one mock survey consistent with
observations. By inspection of Figure \ref{fig:landscape} we see that
the halo number abundance is able to constraint the minimum mass
$M_{\rm min}$ to a narrow range, while $M_{\rm max}$ and $f_{\rm occ}$
remain largely unconstrained.

In Figure \ref{fig:landscape} there are three regions of parameter
space that can be clearly distinguished. The first corresponds to
models where the minimum mass is high $M_{\rm min}>
10^{11.5}\hMsun$.  For these models the number density of LAEs is too low
to be compatible with observations.

The second region corresponds to an intermediate range for the minimum
mass $10^{10.5}\hMsun < M_{\rm min}< 10^{11.5}\hMsun$ where,
regardless of the value of the maximum mass $M_{\rm max}$, it is
possible to tune the occupation fraction $f_{\rm occ}$ to bring some
of the mock surveys into good agreement with observations. In this
region we find two extreme kinds of models. One extreme are models
with a narrow mass interval $\Delta M\equiv \log_{10} M_{\rm max}
- \log_{10}M_{\rm  min} <1.0$ dex. The other extreme are models with a
large mass interval $\Delta M>1.0$ dex going up to the maximum halo
mass present in the simulation at that redshift, with $\Delta M = 2.5$
dex in some cases.

@@

Without any additional information our method allows us to infer that
most of the successful models are found in the second and third regions of
parameter space. This result was expected from halo
abundance calculations shown in Figure \ref{fig:halos} and discussed
in the previous subsection.

@@

%\begin{figure}
%\begin{center}
%\includegraphics[width=0.95\linewidth,angle=0]{Fig4_match_P5.pdf}
%\end{center} 
%\caption{ Number of models with at least $N_{\rm high-P}$ mock surveys
%  consistent with the observed surface number density
%  distribution in terms of the KS-test values $P>0.05$. Only $\sim
%  100$ models have all their mocks consistent with observations. 
%  \label{fig:high_success_rate}.}  
%\end{figure}

@@

@@

In the previous sub-section we use a conservative criterion of
agreement by selecting the models that had at least one mock survey
with $P>0.05$ in the KS test. Now we turn to a more strict selection
by requiring all the 15 constructed mock surveys to be consistent with
observations.

%Figure \ref{fig:high_success_rate} shows the number of models that
%have at least $N_{\rm high-P}$ mocks with $P>0.05$. For each model
%${\mathcal M}$ there are 15 different mock surveys.  We discard the
%models that have 14 mocks or less consistent with observations.

With
this cut we find $\sim 100$ models with all the 15 mock survey
realizations with $P>0.05$.  This cut represents a reduction by a
factor of $\sim 6$ with respect to the total number of models with at
least one consistent mock.

Figure \ref{fig:restriction_mock} presents the locii of these models
in the parameter space $M_{\rm min}-M_{\rm max}$ and $M_{\rm
  min}-f_{\rm occ}$. With this constraints the number of consistent
models with  $10.5 < \log_{10}M_{\rm min}< 11.0$ are reduced. This 
corresponds to the regions in the parameter space in
Figure \ref{fig:landscape} that already had a low number of 
consistent mock surveys. On the other hand, from the right panel in
Figure \ref{fig:restriction_mock} one can see that the favored
occupation fraction  values are still in a range $0\leq f_{\rm
  occ}\leq 1.0$.

We conclude that conditions on the number density statistics, even if they are
very strict, only put strong constraints on the minimum mass $M_{\rm
  min}$ but not on the other two parameter of the model $M_{\rm max}$
and $f_{\rm occ}$.

@@

Figure \ref{fig:correlation_parameters} shows the main results for the
best estimates of $\theta_{0}-\beta$ in the parameterization of the
ACF. Blue circles represent the results from the mocks and the crosses
from observations. In the left panel Figure \ref{fig:correlation_parameters} we see
how the observational ACF measured by \cite{Hayashino2004} 
can be successful in reducing the total number of possible models. The
right panel shows the results for fields in the mock catalog with an
average number density compared against the fit of the data
reported by \cite{Ouchi2010}. In this case we find that there are
large uncertainties coming from the observational data, and only a reduced
fraction of the models can be discarded.

@@

\begin{figure}
\begin{center}
\includegraphics[width=1.0\linewidth,angle=0]{power_law_correlation_meanden.pdf} 
\end{center}
\caption{Values for the free parameters $\theta_{0}$ and $\beta$
in the fitting formula (Eq. \ref{eq:fitting}) for the angular
correlation function. Blue dots correspond to simulations and the
green cross to observations by \citet{Ouchi2010} (right). The error
bars in the theoretical data correspond to the quadratic average of
the fitting errors for each mock survey 
\label{fig:correlation_parameters}}
\end{figure}

@@

\cite{Hayes2010} showed that the luminosity function for
LAEs at $z=2.2$ is consistent with the escape fraction being constant
for every galaxy regardless of its luminosity. From this results they
derive that almost $90\%$ of the star forming galaxies emit insufficient
Lyman $\alpha$ to be detected, effectively setting the occupation
fraction to be $f_{\rm occ}=0.10$. \cite{Dijkstra2013} use similar
methods to find an average \ly escape fraction in high-z galaxies,
they conclude that there is a range in \ly luminosity where the escape
fraction they derive at $z\sim 3.0$, $f_{\rm esc}=(17\pm 5)\%$, could
be interpreted as an occupation fraction $f_{\rm occ}\sim 0.2$. We
consider a success that our model indeed includes such low occupation
fraction models. The observational evidence from galaxy surveys seem
to disfavor the second family of models with high occupation
fractions.

@@

Observational evidence based on the ACF inferred from photometric
measurements in the Extended Chandra Deep Field South has shown that
the median dark matter masses of halos hosting LAEs is
$\log_{10}M_{\rm  med}=10.9^{+0.5}_{-0.9}$\Msun, with a corresponding
occupation fraction of $1-10\%$  \citep{Gawiser07}.  \cite{Ouchi2010}
presents analysis of LAE observations in the redshift interval
$3.1<z<7.0$ and at $z=3.1$ They quote an average mass for the host
dark matter halos of $M_{h}=2.9^{+24.0}_{-2.9}\times 10^{10}$ \hMsun
with a corresponding duty cycle of $0.008\pm 0.03$.
